---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AddFonts

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/gnoblet/AddFonts/workflows/R-CMD-check/badge.svg)](https://github.com/gnoblet/AddFonts/actions)
<!-- badges: end -->

AddFonts provides an easy way to download and use fonts from various providers in R graphics. Currently supports [Bunny Fonts](https://fonts.bunny.net/), a privacy-focused, GDPR-compliant alternative to Google Fonts.

## Features

- 🔒 **Privacy-focused**: Use Bunny Fonts instead of Google Fonts
- 📦 **Simple API**: Similar to `showtext::font_add_google()`
- 💾 **Smart caching**: Downloads fonts once, uses them everywhere
- 🔍 **Easy discovery**: Search for fonts by name or category
- 🎨 **Full support**: Regular, bold, italic, and bold-italic variants

## Installation

You can install the development version of AddFonts from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("gnoblet/AddFonts")
```

## Quick Start

```{r example, eval = FALSE}
library(AddFonts)
library(showtext)

# Add a font from Bunny Fonts
add_font_bunny("open-sans")

# Enable showtext for graphics
showtext_auto()

# Use the font in your plots
plot(1:10, main = "Using Open Sans from Bunny Fonts!")
```

### With ggplot2

```{r ggplot2-example, eval = FALSE}
library(ggplot2)
library(AddFonts)
library(showtext)

# Add custom fonts
add_font_bunny("roboto", family = "Roboto")
add_font_bunny("merriweather", family = "Merriweather")

showtext_auto()

ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    title = "Motor Trend Car Road Tests",
    subtitle = "Relationship between weight and fuel efficiency"
  ) +
  theme_minimal(base_family = "Roboto") +
  theme(
    plot.title = element_text(family = "Merriweather", face = "bold", size = 16)
  )
```

## Searching for Fonts

Find fonts by name or browse by category:

```{r search-example, eval = FALSE}
# Search for fonts by name
font_search_bunny("roboto")

# List all sans-serif fonts
fonts <- font_list_bunny()
sans_fonts <- fonts[fonts$category == "sans-serif", ]
head(sans_fonts)

# Browse all available categories
table(fonts$category)
```

## Advanced Usage

### Custom Font Weights and Variants

```{r advanced-example, eval = FALSE}
# Use a lighter weight for regular text
add_font_bunny(
  "roboto",
  family = "RobotoLight",
  regular.wt = 300,
  bold.wt = 500
)

# Skip italic variants if not needed
add_font_bunny(
  "source-code-pro",
  family = "SourceCodePro",
  italic = FALSE
)

# Specify a different character subset
add_font_bunny(
  "noto-sans",
  family = "NotoSans",
  subset = "latin-ext"
)
```

### Custom Cache Directory

By default, fonts are cached in a platform-appropriate directory. You can customize this:

```{r cache-example, eval = FALSE}
# Check current cache directory
get_font_cache_dir()

# Use a custom cache location
add_font_bunny("open-sans", cache_dir = "~/my-fonts")

# Clear the font cache
clear_font_cache_dir()
```

## Font Database

AddFonts includes a bundled database of **1,700+ fonts** from Bunny Fonts, updated monthly via GitHub Actions. This means:

- ✅ No API calls needed for browsing
- ✅ Fast font search and listing
- ✅ Works offline (after initial font download)
- ✅ Automatically stays up-to-date

The database is regenerated monthly using a Python script:

```bash
# Update the font database manually
uv run scripts/create_bunny_font_db.py
```

## Comparison with Other Packages

| Feature | AddFonts | showtext | systemfonts |
|---------|----------|----------|-------------|
| Google Fonts | ❌ | ✅ | ❌ |
| Bunny Fonts (Privacy) | ✅ | ❌ | ❌ |
| Local fonts | ➡️ Use showtext | ✅ | ✅ |
| Font caching | ✅ | Partial | N/A |
| Easy search | ✅ | ❌ | ❌ |
| Offline catalog | ✅ | ❌ | N/A |

AddFonts is designed to complement `showtext` by providing an easy way to add fonts from privacy-focused providers.

## Why Bunny Fonts?

[Bunny Fonts](https://fonts.bunny.net/) is an open-source, privacy-first alternative to Google Fonts:

- 🔒 **GDPR compliant**: No tracking, no data collection
- 🚀 **Fast CDN**: Globally distributed network
- 🆓 **Free & open**: No accounts or API keys needed
- 🎨 **Full coverage**: Includes most popular Google Fonts

Perfect for users in the EU or anyone concerned about privacy.

## How It Works

1. **Search**: Browse the bundled font database (updated monthly)
2. **Download**: Fetch font files from Bunny's CDN (cached locally)
3. **Register**: Add fonts to R via `sysfonts`
4. **Use**: Enable with `showtext_auto()` and use in any graphics device

```{r workflow, echo = FALSE, eval = FALSE}
# Simplified workflow diagram
# Search fonts → Download & cache → Register → Use in plots
```

## Contributing

Contributions are welcome! Ideas for future development:

- [ ] Support for additional font providers (Fontsource, etc.)
- [ ] Shiny app for interactive font browsing
- [ ] Font preview function
- [ ] Automatic font pairing suggestions
- [ ] Support for variable fonts

## Related Packages

- [showtext](https://github.com/yixuan/showtext): Using fonts in R graphics
- [sysfonts](https://github.com/yixuan/sysfonts): Loading fonts into R
- [systemfonts](https://github.com/r-lib/systemfonts): System font management
- [extrafont](https://github.com/wch/extrafont): TrueType font support

## License

GPL (>= 3)

## Acknowledgments

- [Bunny Fonts](https://fonts.bunny.net/) for providing a privacy-focused font CDN
- [showtext](https://github.com/yixuan/showtext) by Yixuan Qiu for the underlying font rendering
- Font database updated automatically via GitHub Actions

---

**Note**: AddFonts is in early development. The API may change. Feedback and contributions welcome!
