---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AddFonts

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/gnoblet/AddFonts/workflows/R-CMD-check/badge.svg)](https://github.com/gnoblet/AddFonts/actions)
[![R-CMD-check](https://github.com/gnoblet/AddFonts/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gnoblet/AddFonts/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/gnoblet/AddFonts/graph/badge.svg)](https://app.codecov.io/gh/gnoblet/AddFonts)
<!-- badges: end -->

AddFonts provides an easy way to download and use fonts from various providers in R graphics. Currently supports [Bunny Fonts](https://fonts.bunny.net/), a privacy-focused, GDPR-compliant alternative to Google Fonts.

## Features

- 🔒 **Privacy-focused**: Use Bunny Fonts instead of Google Fonts
- 📦 **Simple**: Similar to `showtext::font_add_google()`
- 💾 **Smart caching**: Downloads fonts once, uses them everywhere
- 🔍 **Easy discovery**: Search for fonts by name or category
- 🎨 **Full support**: Regular, bold, italic, and bold-italic variants

## Installation

You can install the development version of AddFonts from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("gnoblet/AddFonts")
```

### System Requirements

AddFonts requires the `woff2_decompress` command-line tool to convert downloaded font files to a format compatible with R graphics. Install it using your system's package manager:

```bash
# macOS (Homebrew)
brew install woff2

# Debian/Ubuntu
sudo apt install woff2

# Fedora/RHEL
sudo dnf install woff2-tools

# Arch Linux
sudo pacman -S woff2

# Windows
# Build from source: https://github.com/google/woff2
```

## Quick Start

```{r example}
library(AddFonts)
library(showtext)

# Add a font from Bunny Fonts (downloads all weights and styles by default)
add_font("open-sans")

# Or use the provider-specific function
add_font_bunny("roboto")

# Enable showtext for graphics
showtext_auto()

# Use the font in your plots
plot(1:10, main = "Using Open Sans from Bunny Fonts!")
```

### With ggplot2

```{r ggplot2-example}
library(ggplot2)
library(AddFonts)
library(showtext)

# Add custom fonts with specific weights
add_font_bunny(
  "Cormorant SC",
  family = "main_font",
  regular_wt = 600,
  bold_wt = 700
)

showtext_auto()

ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(
    title = "Motor Trend Car Road Tests",
    subtitle = "Relationship between weight and fuel efficiency"
  ) +
  theme_minimal(base_family = "main_font") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 14)
  )
```

## Searching for Fonts

Find fonts by name or browse by category:

```{r search-example, eval=false}
# Search for fonts by name (uses default provider: Bunny)
font_search("roboto")

# Or use provider-specific function
font_search_bunny("roboto")

# List all available fonts
fonts <- font_list()
head(fonts)

# List all sans-serif fonts
sans_fonts <- font_search(category = "sans-serif")

# Browse all available categories
table(fonts$category)
```

## Advanced Usage

### Custom Font Weights and Styles

```{r advanced-example, eval=false}
# Download all available weights and styles (default)
add_font("roboto")

# Download only specific weights
add_font("roboto", wt = c(300, 400, 700))

# Download only normal (non-italic) styles
add_font("roboto", wt = c(400, 700), styles = "normal")

# Download only italic styles with specific weights
add_font("merriweather", wt = c(400, 700), styles = "italic")

# Download a single weight
add_font("source-code-pro", wt = 400, styles = "normal")

# Specify a different character subset
add_font("noto-sans", subset = "latin-ext")

# You can also use provider-specific functions
add_font_bunny("inter", wt = c(300, 500), styles = "normal")
```

### Custom Cache Directory

By default, fonts are cached in a platform-appropriate directory. You can customize this:

```{r cache-example, eval=false}
# Check current cache directory
get_font_cache_dir()

# Use a custom cache location
add_font_bunny("open-sans", cache_dir = "~/my-fonts")

# Clear the font cache
clear_font_cache_dir()
```

## Font Database

AddFonts includes a bundled database of **1,700+ fonts** from Bunny Fonts, updated monthly via GitHub Actions. This means:

- ✅ No API calls needed for browsing
- ✅ Fast font search and listing
- ✅ Works offline (after initial font download)
- ✅ Automatically stays up-to-date

The database is regenerated monthly using a Python script:

```bash
# Update the font database manually
uv run scripts/create_bunny_font_db.py
```

## Why Bunny Fonts?

[Bunny Fonts](https://fonts.bunny.net/) is an open-source, privacy-first alternative to Google Fonts:

- 🔒 **GDPR compliant**: No tracking, no data collection
- 🚀 **Fast CDN**: Globally distributed network
- 🆓 **Free & open**: No accounts or API keys needed
- 🎨 **Full coverage**: Includes most popular Google Fonts

Perfect for users in the EU or anyone concerned about privacy.

## How It Works

1. **Search**: Browse the bundled font database (updated monthly)
2. **Download**: Fetch WOFF2 font files from Bunny's CDN (cached locally)
3. **Convert**: Automatically convert WOFF2 to TTF format for R compatibility
4. **Register**: Add fonts to R via `sysfonts`
5. **Use**: Enable with `showtext_auto()` and use in any graphics device

**Why the conversion?** Bunny Fonts provides fonts in WOFF2 format (optimized for web use), but R's `sysfonts` doesn't support WOFF2 out-of-the-box. AddFonts automatically converts downloaded fonts to TTF format for maximum compatibility.

**Performance**: Conversion happens only once per font file. Subsequent uses of the same font load instantly from the TTF cache.

## Related Packages

- [showtext](https://github.com/yixuan/showtext): Using fonts in R graphics
- [sysfonts](https://github.com/yixuan/sysfonts): Loading fonts into R
- [systemfonts](https://github.com/r-lib/systemfonts): System font management
- [extrafont](https://github.com/wch/extrafont): TrueType font support

## License

GPL (>= 3)
---

**Note**: AddFonts is in early development. Everything may change. Feedback and contributions welcome!
